---
title: "Data Incubator Project Proposal"
author: "Robert Handsfield"
date: "04/10/2015"
output: html_document
---

Written in R markdown

Correlate real estate prices to natural disasters.


```{r}
library("data.table");
library("dplyr");
library("ggplot2");
library("reshape");
```


The weather dataset, codebooks, and supplementary information may be found at the following links.
This analysis was performed on the NOAA Storm Data Set obtained on **February 2^nd^, 2015**.

* NOAA [Storm Data Set](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)
* Storm Data [Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)
* NCDC Storm Events [FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)
* Proprietary [Codebook](https://ire.org/nicar/database-library/databases/storm-events/) from IRE
* [Revisions](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf) to the data set

```{r downloadSource, echo=TRUE, cache=TRUE, eval=FALSE}
# Download the file
fileUrl <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2";

if ( !("Stormdata.csv.bz2" %in% list.files()) ) {
	download.file(fileUrl, destfile = "./Stormdata.csv.bz2", method = "curl" );
}
```  

```{r eval=FALSE}
file <- "../datasciencecoursera/5-reproducibleResearch/courseProject2/Stormdata.csv.bz2";
tmp <- bzfile(file);
dfStorm <- read.csv(tmp);
```

```{r}
dt <- data.table(dfStorm);

dt$BGN_DATE <- as.Date(dt$BGN_DATE, format= "%m/%d/%Y");

#dt$BGN_DATE[1] < 2003;

nrow( dt[BGN_DATE < 1980,] );

# subset events that cause housing damage: tornado, flood, hurricane, wildfire
storm.kind <- c(unique(dt$EVTYPE)[ grep("WILDFIRE", unique(dt$EVTYPE))],
		unique(dt$EVTYPE)[ grep("TORNADO", unique(dt$EVTYPE))],
		unique(dt$EVTYPE)[ grep("HURRICANE", unique(dt$EVTYPE))]
		);

```

```{r}
# subset events that cause housing damage: tornado, flood, hurricane, wildfire

# create list of storm events
storm.kind <- c(grep("FIRE", unique(dt$EVTYPE)),
		grep("TORNADO", unique(dt$EVTYPE)),
		grep("HURRICANE", unique(dt$EVTYPE)),
		grep("FLOOD", unique(dt$EVTYPE))
		);

# convert indexes to names (factors)
storm.kind <- unique(dt$EVTYPE)[storm.kind];

dt <- dt[EVTYPE %in% storm.kind,];
dt$EVTYPE <- factor(dt$EVTYPE); # remove extraneous factors

# subset continental U.S.
state.list <- c('AL', 'AZ', 'AR', 'CA', 'CO', 'CT', 'DE', 'DC', 'FL', 'GA', 'ID', 'IL', 'IN', 'IA', 'KS', 'KY', 'LA', 'ME', 'MD', 'MA', 'MI', 'MN', 'MS', 'MO', 'MT', 'NE', 'NV', 'NH', 'NJ', 'NM', 'NY', 'NC', 'ND', 'OH', 'OK', 'OR', 'PA', 'RI', 'SC', 'SD', 'TN', 'TX', 'UT', 'VT', 'VA', 'WA', 'WV', 'WI', 'WY');

dt <- dt[STATE %in% state.list];
dt$STATE <- factor(dt$STATE); # remove extraneous factors

# real estate data available from 2009
dt <- dt[BGN_DATE > 2008];

# create state.county factor
dt$state.county <- paste(dt$STATE, dt$COUNTYNAME, sep=".");

tbl <- table(dt$state.county)

which(tbl == max(tbl));

sort(tbl, decreasing=TRUE )[1:100];

top100 <- names( sort(tbl, decreasing=TRUE )[1:100]);
```


Harris County TX (Houston) is 4^th^ on the list of counties with highest incidence of fires, floods, tornadoes, or hurricanes.

```{r}
# subset to harris county
dt.harris <- dt[state.county == "TX.HARRIS", ];

# subset to 2009
dt.harris.09 <- dt.harris[BGN_DATE > "2009-10-01",];
event <- dt.harris.09$BGN_DATE;

ggplot(dt.harris) + aes(x=BGN_DATE) + geom_bar(stat="bin") + ggtitle("Frequency of Natural Disasters in Harris County, TX");
```



---------------------------------------------

Obtain real estate statistics for Harris county, TX, via the Trulia API:

http://api.trulia.com/webservices.php?library=TruliaStats&function=getCountyStats&county=Harris&state=TX&startDate=1950-01-03&endDate=2011-11-03&apikey=wwccxrp5983jhuaxj278fpbt

```{r}
# real estate data parsed and saved to 'harris2.csv'

harris <- data.table(read.csv("data/harris2.csv")[1:4]);
harris$week.ending <- as.Date(harris$week.ending, format="%Y-%m-%d");

re.harris <- data.frame(harris);

tmp1 <- data.frame(re.harris$week.ending); 
tmp1$value <-  re.harris$num.props;
tmp1$measurement <- "properties listed";

tmp2 <- data.frame(re.harris$week.ending);
tmp2$value <-  re.harris$med.price;
tmp2$measurement <- "median price";

tmp3 <- data.frame(re.harris$week.ending);
tmp3$value <-  re.harris$avg.price;
tmp3$measurement <- "mean price";

re.harris2 <- rbind(tmp1, tmp2, tmp3);
names(re.harris2)[1] <- "week.ending";

```

```{r}

ggplot(re.harris2) + aes(x=week.ending, y=value, color=measurement) + geom_line() + 
	scale_fill_continuous(guide = "legend") + 
	geom_point(aes(x = event[1], y=3e05), color="black") + 
	geom_point(aes(x = event[2], y=3e05), color="black") +
	geom_point(aes(x = event[3], y=3e05), color="black") +
	geom_point(aes(x = event[4], y=3e05), color="black") +
	geom_point(aes(x = event[5], y=3e05), color="black") +
	geom_point(aes(x = event[6], y=3e05), color="black") +
	geom_point(aes(x = event[7], y=3e05), color="black") +
	geom_point(aes(x = event[8], y=3e05), color="black") +
	geom_point(aes(x = event[9], y=3e05), color="black") +
	geom_point(aes(x = event[10], y=3e05), color="black") + 
	ggtitle("Real Estate Prices and Natural Disasters for Harris County, TX");

```

By week, the number of listed properties is shown in green, the median listing price in red, and the average listing price in blue. Disasters- floods, tornadoes, etc, are plotted as black dots on the date they occurred.

It's unclear whether this small sample shows a correlation.  The topic is worthy of further investigation.